Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling

### 2025-01-03 hybrid.tex
Edits before format review

This chapter starts with the $S_N$ method implementation required for the hybrid $S_N$-diffusion
method. I implemented both $S_N$ and hybrid methods using \gls{FEM} numerical solver capabilities
available through Moltres \cite{lindsay_moltres_2017} and the \gls{MOOSE} framework. In this
section, the derivation for the \gls{SAAF} formulation of the $S_N$ method and its
implementation in Moltres follows the derivation developed by Wang et al.\
\cite{wang_diffusion_2014, wang_rattlesnake_2018}.

@@

Continuing from the introduction to neutronics methods in Section \ref{sec:summary-nts-mtds},
the time-dependent, multigroup neutron transport equation defined on the 3-D spatial domain
$\mathcal{D}$ and 2-D unit sphere angular domain $\mathcal{S}$ is:
  \frac{\partial}{\partial t}\left[\frac{\Psi_g(\vec{r},\hat{\Omega},t)}{v_g}\right] +
  \hat{\Omega}\cdot\nabla\Psi_g(\vec{r},\hat{\Omega},t) + \Sigma_{t,g}
  \Psi_g(\vec{r},\hat{\Omega},t) =
  \sum^G_{g'=1}\int_\mathcal{S} \Sigma_s^{g'\rightarrow g}(\hat{\Omega}'\rightarrow\hat{\Omega})
  \Psi_{g'}(\vec{r},\hat{\Omega}',t)d\hat{\Omega}' \\
  + \frac{1}{4\pi}\chi_{p,g}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'}(\vec{r},t)
  + \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,g}
  \lambda_i C_i(\vec{r},t) + Q^{\text{ext}}_g(\vec{r},\hat{\Omega})
  \label{eq:mg-nte}
with the boundary conditions
  \Psi_g(\vec{r},\hat{\Omega}) = \Psi^\text{inc}_g(\vec{r},\hat{\Omega}) +
  \alpha^s_g\Psi_g(\vec{r},\hat{\Omega}_r)
  \mbox{ on } \vec{r} \in \partial\mathcal{D} \mbox{ and } \hat{\Omega}\cdot\hat{n}_b < 0,
  \begin{align*}
    \chi_{p,g} &= \mbox{prompt fission neutron spectrum in group $g$,} \\
    \beta &= \sum^I_{i=1} \beta_i = \mbox{total delayed neutron fraction,} \\
    \chi_{d,g} &= \mbox{delayed fission neutron spectrum in group $g$,} \\
    \lambda_i &= \mbox{decay constant of precursor group $i$,} \\
    C_i &= \mbox{delayed neutron precursor concentration for group $i$,} \\
    \Psi^\text{inc}_g &= \mbox{incident surface source in group $g$,} \\
    \alpha^s_g &= \mbox{specular reflectivity on }\partial \mathcal{D} \mbox{ for group }g, \\
    \hat{\Omega}_r &= \hat{\Omega}-2(\hat{\Omega}\cdot \hat{n}_b)\hat{n}_b, \\
    \hat{n}_b &= \mbox{outward unit normal vector on the boundary.}
  \end{align*}
In order to introduce operators and facilitate subsequent mathematical derivations, we will define
the following vector forms:
  \bm{\Psi} \equiv
    \Psi_1 \\
    \Psi_2 \\
    \Psi_G
  \bm{\Phi} \equiv \int_S \bm{\Psi}d\hat{\Omega} \equiv
    \phi_1 \\
    \phi_2 \\
    \phi_G
  \bm{\frac{\Psi}{v}} \equiv
    \frac{\Psi_1}{v_1} \\
    \frac{\Psi_2}{v_2} \\
    \frac{\Psi_G}{v_G}
  \bm{C} \equiv
    C_1 \\
    C_2 \\
    C_G
  \bm{Q}^{\text{ext}} \equiv
    Q^\text{ext}_1 \\
    Q^\text{ext}_2 \\
    Q^\text{ext}_G
%  \bm{Q}_f = \frac{1}{4\pi}\bm{Q}_{f,0} =
%  \begin{bmatrix}
%    \frac{1}{4\pi}\chi_{p,1}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
%    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,1} \lambda_i C_i \\
%    \frac{1}{4\pi}\chi_{p,2}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
%    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,2} \lambda_i C_i \\
%    \vdots \\
%    \frac{1}{4\pi}\chi_{p,G}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
%    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,G} \lambda_i C_i
%  \end{bmatrix},
%  \bm{Q} = \bm{Q}_f + \bm{Q}^\text{ext}. \nonumber
We also define the following operators:
  \mathbb{L}_1\bm{\Psi} \equiv
    \hat{\Omega}\cdot\nabla\Psi_1 \\
    \hat{\Omega}\cdot\nabla\Psi_2 \\
    \hat{\Omega}\cdot\nabla\Psi_G \\
  \mathbb{L}_2\bm{\Psi} \equiv
    \Sigma_{t,1}\Psi_1 \\
    \Sigma_{t,2}\Psi_2 \\
    \Sigma_{t,G}\Psi_G
  \mathbb{L}\bm{\Psi} \equiv \mathbb{L}_1\bm{\Psi} + \mathbb{L}_2\bm{\Psi}, \nonumber \\
  \mathbb{S}\bm{\Psi} \equiv
    \sum^G_{g'=1}\int_S \Sigma_s^{g'\rightarrow 1}\Psi_{g'}d\hat{\Omega} \\
    \sum^G_{g'=2}\int_S \Sigma_s^{g'\rightarrow 2}\Psi_{g'}d\hat{\Omega} \\
    \sum^G_{g'=G}\int_S \Sigma_s^{g'\rightarrow G}\Psi_{g'}d\hat{\Omega}
  \mathbb{B}\bm{\Psi} \equiv
    \alpha^s_1\Psi_1(\hat{\Omega}_r) \\
    \alpha^s_2\Psi_2(\hat{\Omega}_r) \\
    \alpha^s_G\Psi_G(\hat{\Omega}_r)
  \mathbb{F}\bm{\Psi} \equiv \frac{1}{4\pi}\mathbb{F}_0\bm{\Psi} \equiv
    \frac{1}{4\pi}\chi_{p,1}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'} \\
    \frac{1}{4\pi}\chi_{p,2}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'} \\
    \frac{1}{4\pi}\chi_{p,G}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}
  \mathbb{C}\bm{C} \equiv \frac{1}{4\pi}\mathbb{C}_0\bm{C} \equiv
    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,1} \lambda_i C_i \\
    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,2} \lambda_i C_i \\
    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,G} \lambda_i C_i
  \end{bmatrix}. \nonumber
Note that the operators apply element-wise multiplication as opposed to the more conventional
matrix multiplication. Eq.\ \ref{eq:mg-nte} can be reexpressed as:
  \frac{\partial}{\partial t} \left(\frac{\bm{\Psi}}{\bm{v}}\right)+\mathbb{L}\bm{\Psi}
  = \mathbb{S}\bm{\Psi} + \bm{Q}, \label{eq:nte-vec}
with the boundary conditions on $\partial\mathcal{D}$
  \bm{\Psi} = \bm{\Psi}^\text{inc} + \mathbb{B}\bm{\Psi},
  \begin{align*}
    \bm{Q} &= \bm{Q}_f + \bm{Q}^\text{ext}, & \\
    \bm{Q}_f &= \mathbb{F}\bm{\Psi} + \mathbb{C}\bm{C}. &
  \end{align*}

@@

Thus, we can define the inner products
%
\begin{gather}
  \left(\bm{a},\bm{b}\right) \equiv \sum^G_{g=1}\int_\mathcal{S}d\hat{\Omega}
  \sum_{e\in\mathbb{T}_h}\int_{\mathcal{D}_e}d\vec{r}\sum_{i\in E_e}a_{g,i}(\hat{\Omega})
  b_i(\vec{r})\sum_{j\in E_e}b_{g,j}(\hat{\Omega})b_j(\vec{r}), \\
  \langle\bm{a},\bm{b}\rangle^{\pm} \equiv \sum^G_{g=1}\sum_{s\in\partial\mathcal{D}}\int_s
  d\vec{r} \int_{mathcal{S}^\pm_{\hat{n}_b}}d\hat{\Omega} |\hat{\Omega}\cdot\hat{n}_b|
  \sum_{i\in E_s}a_{g,i}(\hat{\Omega})b_i(\vec{r})\sum_{j\in E_s}b_{g,j}(\hat{\Omega})b_j(\vec{r}),
  \shortintertext{where}
  \begin{align*}
    E_e &= \mbox{index set of the local basis functions in element $e$,} \\
    E_s &= \mbox{index set of the local basis functions in element side $s$,}
  \end{align*}
\end{gather}
%
to apply spatial discretization on Eq.\ \ref{eq:saaf} by replacing the preexisting inner product
notations defined in Eqs.\ \ref{eq:weak-domain} and \ref{eq:weak-boundary}.

@@

Finally, we present the expanded forms of each term in the \gls{SAAF} formulation of the multigroup
$S_N$ neutron transport equations. With the angular discretization scheme laid out, we define the
inner products
%
\begin{gather}
  \left(\bm{a},\bm{b}\right)_\mathcal{D} \equiv \sum^G_{g=1} \int_\mathcal{D}d\vec{r}
  \sum_{i\in E_e}a_{g,i}b_i(\vec{r})\sum_{j\in E_e}b_{g,j}b_j(\vec{r}), \\
  \left(\bm{a},\bm{b}\right)_{\partial\mathcal{D}} \equiv \sum^G_{g=1}
  \sum_{s\in\partial\mathcal{D}}\int_s d\vec{r}\sum_{i\in E_s}a_{g,i}b_i(\vec{r})\sum_{j\in E_s}
  b_{g,j}b_j(\vec{r}),
\end{gather}
%
involving integrals over the spatial domain to explicitly show the $S_N$ angular discretization.

\noindent Time derivative term:
%
\begin{gather}
  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,
  \frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right)\right) =
  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\Psi^*_{g,d},
  \frac{\Psi_{g,d}}{v_g}\right)_\mathcal{D} \label{eq:time-derivative}
\end{gather}
%
Streaming term:
%
\begin{gather}
  \left(\mathbb{L}_1\bm{\Psi}^*,
  \left(\bm{\tau}\mathbb{L}_1-\mathbb{I}+\bm{\tau}\mathbb{L}_2\right)\bm{\Psi}\right) =
  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},\tau_g\hat{\Omega}
  \cdot\nabla\Psi_{g,d}-(1-\tau_g\Sigma_{t,g})\Psi_{g,d}\right)_\mathcal{D}
\end{gather}
%
Collision term:
%
\begin{gather}
  \left(\mathbb{L}_2\bm{\Psi}^*,\bm{\Psi}\right) =
  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},\Sigma_{t,g}\Psi_{g,d}\right)_\mathcal{D}
\end{gather}
%
Scattering term:
%
\begin{gather}
  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{S}\bm{\Psi}\right) =
  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
  \sum^G_{g'=1}\sum^L_{l=0}\Sigma^{g'\rightarrow g}_{s,l}\sum^l_{m=-l}
  \frac{2l+1}{w}Y_{l,m}(\hat{\Omega}_d)\phi_{g',l,m}\right)_\mathcal{D}
\end{gather}
%
Prompt fission source term:
%
\begin{gather}
  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{F}\bm{\Psi}\right) =
  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
  \frac{1}{w}\bar{\chi}_g\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}\right)_\mathcal{D}
\end{gather}
%
where $\bar{\chi}_g=\chi_{p,g}\left(1-\beta\right)$ for time-dependent problems with \glspl{DNP}
and $\bar{\chi}_g=\chi_{g}$ for steady-state problems without \glspl{DNP}.

@@

Regardless of the differences in implementations amongst the methods discussed here,
diffusion corrections applied through Eq.\ \ref{eq:saaf} were
found to be effective in enabling the neutron diffusion method to accurately reproduce reference
flux solutions \cite{gross_comprehensive_2023, pounders_diffusion_2009}.
However, a significant challenge for diffusion corrections
involves their resolution near neutron flux peaks. As observed in Eq.\ \ref{eq:svdc}, the neutron
current and flux gradient values do not necessarily reach zero at the same points in space,
resulting in diffusion coefficient values tending to positive or negative infinity when the
flux gradient is close to zero. Larger mesh sizes may help avoid this issue at the cost of
worsening flux accuracy in regions with
steep, non-linear fluxes, such as near control rods. For the Ronen method, Gross et al.\
\cite{gross_comprehensive_2023} applied a numerical fix by avoiding correction
calculations wherever the flux gradient values fall below a certain threshold.

@@

equations. For this work, we will adopt drift term derivations developed by Wang et al.\
\cite{wang_diffusion_2014, wang_rattlesnake_2018}. We start by integrating Eq.\ \ref{eq:saaf-vt}

@@

In this discussion thus far, I have yet to establish the cutoff boundary for $V_1'$ within $V_1$
opposite to $\partial V_1$. A natural/intuitive criterion for the location of the cutoff boundary
would be where the components of the drift correction variable $\vec{D}_g$ become zero (i.e.,
where the components change signs). The reasons for this criterion are twofold. Firstly, at
points where the $\vec{D}_g$ components are zero, the flux is approximately isotropic along the
axes corresponding to the components because value of $\vec{D}_g$ represents how much correction a
standard neutron diffusion model requires. Secondly, this choice preserves the smoothness of the
neutron flux and flux gradient.
%
\begin{figure}[htb!]
	\centering
	\includegraphics[width=.75\columnwidth]{case-3b-geometry}
	\caption{1-D geometry for Case 3b.}
	\label{fig:3b-geometry}
\end{figure}

Here, I use Case 3b of the 1-D test cases to be covered in Chapter \ref{chap:msre}
to illustrate the buffer region and the behavior of $\vec{D}_g$. Case 3b is a 1-D slab problem
consisting of control rod, air, fuel-graphite repeating lattice, and reactor vessel regions as
shown in Figure \ref{fig:3b-geometry}. The fuel regions are neutron-multiplying media containing
$^\text{235}$U fissile material. I took material compositions of each region from the \gls{MSRE}
reference design and the start-up fuel composition \cite{fratoni_molten_2020}.
Case 3b has reflecting boundary conditions on the left boundary and vacuum boundary conditions on
the right boundary.
%
\begin{figure}[htb!]
    \centering
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\textwidth]{case-3b-group-1-drift}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.49\textwidth}
        \centering
        \includegraphics[width=\textwidth]{case-3b-group-8-drift}
    \end{subfigure}
    \caption{The reference and hybrid drift ($\vec{D}_g$) distributions for group 1 and 8 calculated
      from $S_8$ and $S_8$-diffusion simulations. The correction subregion $V_1$ spans $x=0$ cm to
      $x=10$ cm.}
    \label{fig:3b-drift-1}
\end{figure}

>>>>

The second-order linear \gls{SAAF} formulation of the $S_N$ neutron transport equations is
compatible with the highly efficient and scalable HYPRE-BoomerAMG preconditioner
\cite{hypre_hypre_2022} available through
the \gls{MOOSE} framework. Implementing the $S_N$ method in Moltres allows it to natively interface
with the existing neutron diffusion solver for nonlinear diffusion acceleration and for the
subsequent hybrid method implementation.
The derivation of the \gls{SAAF} formulation of the $S_N$ method and its
implementation in Moltres follows the derivation developed by Wang et al.\
\cite{wang_diffusion_2014, wang_rattlesnake_2018}.

@@

Continuing from the introduction to neutronics methods in Section \ref{sec:summary-nts-mtds},
the time-dependent, multigroup neutron transport equation defined on the 3-D spatial domain
$\mathcal{D}$ and 2-D unit sphere angular domain $\mathcal{S}$ is:
%
\begin{multline}
  \frac{\partial}{\partial t}\left[\frac{\Psi_g(\vec{r},\hat{\Omega},t)}{v_g(\vec{r},t)}\right] +
  \hat{\Omega}\cdot\nabla\Psi_g(\vec{r},\hat{\Omega},t) + \Sigma_{t,g}(\vec{r},t)
  \Psi_g(\vec{r},\hat{\Omega},t) \\
  = \sum^G_{g'=1}\int_\mathcal{S} \Sigma_s^{g'\rightarrow g}(\vec{r}, \hat{\Omega}'\rightarrow\hat{\Omega}, t)
  \Psi_{g'}(\vec{r},\hat{\Omega}',t)d\hat{\Omega}' \\
  + \frac{1}{4\pi}\chi_{p,g}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'}(\vec{r},t) \phi_{g'}(\vec{r},t)
  + \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,g}
  \lambda_i C_i(\vec{r},t) + Q^{\text{ext}}_g(\vec{r},\hat{\Omega})
  \label{eq:mg-nte}
\end{multline}
%
with the boundary conditions
%
\begin{gather}
  \Psi_g(\vec{r},\hat{\Omega}) = \Psi^\text{inc}_g(\vec{r},\hat{\Omega}) +
  \alpha^s_g\Psi_g(\vec{r},\hat{\Omega}_r)
  \mbox{ on } \vec{r} \in \partial\mathcal{D} \mbox{ and } \hat{\Omega}\cdot\hat{n}_b < 0,
  \shortintertext{where}
  \begin{align*}
    \chi_{p,g}(\vec{r},t) &= \mbox{prompt fission neutron spectrum in group $g$,} \\
    \beta(\vec{r},t) &= \sum^I_{i=1} \beta_i(\vec{r},t) = \mbox{total delayed neutron fraction,} \\
    \chi_{d,g}(\vec{r},t) &= \mbox{delayed fission neutron spectrum in group $g$,} \\
    \lambda_i(\vec{r},t) &= \mbox{decay constant of precursor group $i$,} \\
    C_i(\vec{r},t) &= \mbox{delayed neutron precursor concentration for group $i$,} \\
    \Psi^\text{inc}_g(\vec{r},t) &= \mbox{incident surface source in group $g$,} \\
    \alpha^s_g &= \mbox{specular reflectivity on }\partial \mathcal{D} \mbox{ for group }g, \\
    \hat{\Omega}_r &= \hat{\Omega}-2(\hat{\Omega}\cdot \hat{n}_b)\hat{n}_b, \\
    \hat{n}_b &= \mbox{outward unit normal vector on the boundary.}
  \end{align*}
\end{gather}
%
The definitions of the remaining variables may be found in Eq.\ \ref{eq:bte}.
We drop the phase space dependence notation from this point on for brevity.

In order to introduce operators and facilitate subsequent mathematical derivations, we will define
the following vector forms:
%
\begin{gather}
  \bm{\Psi} \equiv
  \begin{bmatrix}
    \Psi_1 \\
    \Psi_2 \\
    \vdots \\
    \Psi_G
  \end{bmatrix},
  \bm{\Phi} \equiv \int_S \bm{\Psi}d\hat{\Omega} \equiv
  \begin{bmatrix}
    \phi_1 \\
    \phi_2 \\
    \vdots \\
    \phi_G
  \end{bmatrix},
  \bm{\frac{\Psi}{v}} \equiv
  \begin{bmatrix}
    \frac{\Psi_1}{v_1} \\
    \frac{\Psi_2}{v_2} \\
    \vdots \\
    \frac{\Psi_G}{v_G}
  \end{bmatrix},
  \bm{C} \equiv
  \begin{bmatrix}
    C_1 \\
    C_2 \\
    \vdots \\
    C_G
  \end{bmatrix},
  \bm{Q}^{\text{ext}} \equiv
  \begin{bmatrix}
    Q^\text{ext}_1 \\
    Q^\text{ext}_2 \\
    \vdots \\
    Q^\text{ext}_G
  \end{bmatrix}, \nonumber
%  \bm{Q}_f = \frac{1}{4\pi}\bm{Q}_{f,0} =
%  \begin{bmatrix}
%    \frac{1}{4\pi}\chi_{p,1}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
%    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,1} \lambda_i C_i \\
%    \frac{1}{4\pi}\chi_{p,2}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
%    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,2} \lambda_i C_i \\
%    \vdots \\
%    \frac{1}{4\pi}\chi_{p,G}(1-\beta)\sum^G_{g'=1} \nu\Sigma_{f,g'} \phi_{g'} +
%    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,G} \lambda_i C_i
%  \end{bmatrix},
%  \bm{Q} = \bm{Q}_f + \bm{Q}^\text{ext}. \nonumber
\end{gather}
%
We also define the following operators:
%
\begin{gather}
  \mathbb{L}_1\bm{\Psi} \equiv
  \begin{bmatrix}
    \hat{\Omega}\cdot\nabla\Psi_1 \\
    \hat{\Omega}\cdot\nabla\Psi_2 \\
    \vdots \\
    \hat{\Omega}\cdot\nabla\Psi_G \\
  \end{bmatrix},
  \mathbb{L}_2\bm{\Psi} \equiv
  \begin{bmatrix}
    \Sigma_{t,1}\Psi_1 \\
    \Sigma_{t,2}\Psi_2 \\
    \vdots \\
    \Sigma_{t,G}\Psi_G
  \end{bmatrix},
  \mathbb{L}\bm{\Psi} \equiv \mathbb{L}_1\bm{\Psi} + \mathbb{L}_2\bm{\Psi}, \nonumber \\
  \mathbb{S}\bm{\Psi} \equiv
  \begin{bmatrix}
    \sum^G_{g'=1}\int_S \Sigma_s^{g'\rightarrow 1}\Psi_{g'}d\hat{\Omega} \\
    \sum^G_{g'=2}\int_S \Sigma_s^{g'\rightarrow 2}\Psi_{g'}d\hat{\Omega} \\
    \vdots \\
    \sum^G_{g'=G}\int_S \Sigma_s^{g'\rightarrow G}\Psi_{g'}d\hat{\Omega}
  \end{bmatrix},
  \mathbb{B}\bm{\Psi} \equiv
  \begin{bmatrix}
    \alpha^s_1\Psi_1(\hat{\Omega}_r) \\
    \alpha^s_2\Psi_2(\hat{\Omega}_r) \\
    \vdots \\
    \alpha^s_G\Psi_G(\hat{\Omega}_r)
  \end{bmatrix}, \nonumber \\
  \mathbb{F}\bm{\Psi} \equiv \frac{1}{4\pi}\mathbb{F}_0\bm{\Psi} \equiv
  \begin{bmatrix}
    \frac{1}{4\pi}\chi_{p,1}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'} \\
    \frac{1}{4\pi}\chi_{p,2}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'} \\
    \vdots \\
    \frac{1}{4\pi}\chi_{p,G}(1-\beta)\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}
  \end{bmatrix},
  \mathbb{C}\bm{C} \equiv \frac{1}{4\pi}\mathbb{C}_0\bm{C} \equiv
  \begin{bmatrix}
    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,1} \lambda_i C_i \\
    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,2} \lambda_i C_i \\
    \vdots \\
    \frac{1}{4\pi}\sum^I_{i=1}\chi_{d,G} \lambda_i C_i
  \end{bmatrix}. \nonumber
\end{gather}
%
Note that the operators apply element-wise multiplication as opposed to the more conventional
matrix multiplication. Eq.\ \ref{eq:mg-nte} can be reexpressed as:
%
\begin{gather}
  \frac{\partial}{\partial t} \left(\frac{\bm{\Psi}}{\bm{v}}\right)+\mathbb{L}\bm{\Psi}
  = \mathbb{S}\bm{\Psi} + \bm{Q}, \label{eq:nte-vec}
\end{gather}
%
with the boundary conditions on $\partial\mathcal{D}$
%
\begin{gather}
  \bm{\Psi} = \bm{\Psi}^\text{inc} + \mathbb{B}\bm{\Psi},
  \shortintertext{where}
  \begin{align*}
    \bm{Q} &= \bm{Q}_f + \bm{Q}^\text{ext}, & \\
    \bm{Q}_f &= \mathbb{F}\bm{\Psi} + \mathbb{C}\bm{C}. &
  \end{align*}
\end{gather}

@@

Thus, we can define the inner products
%
\begin{gather}
  \left(\bm{a},\bm{b}\right) \equiv \sum^G_{g=1}\int_\mathcal{S}d\hat{\Omega}
  \sum_{e\in\mathbb{T}_h}\int_{\mathcal{D}_e}d\vec{r}\sum_{i\in E_e}a_{g,i}(\hat{\Omega})
  b_i(\vec{r})\sum_{j\in E_e}b_{g,j}(\hat{\Omega})b_j(\vec{r}), \\
  \langle\bm{a},\bm{b}\rangle^{\pm} \equiv \sum^G_{g=1}\sum_{s\in\partial\mathcal{D}}\int_s
  d\vec{r} \int_{\mathcal{S}^\pm_{\hat{n}_b}}d\hat{\Omega} |\hat{\Omega}\cdot\hat{n}_b|
  \sum_{i\in E_s}a_{g,i}(\hat{\Omega})b_i(\vec{r})\sum_{j\in E_s}b_{g,j}(\hat{\Omega})b_j(\vec{r}),
  \shortintertext{where}
  \begin{align*}
    E_e &= \mbox{index set of the local basis functions in element $e$,} \\
    E_s &= \mbox{index set of the local basis functions in element side $s$,}
  \end{align*}
\end{gather}
%
to apply spatial discretization on Eq.\ \ref{eq:saaf} by replacing the preexisting inner product
notations defined in Eqs.\ \ref{eq:weak-domain} and \ref{eq:weak-boundary}.

@@

Finally, we present the expanded forms of each term in the \gls{SAAF} formulation of the multigroup
$S_N$ neutron transport equations. With the angular discretization scheme laid out, we define the
inner products
%
\begin{gather}
  \left(\bm{a},\bm{b}\right)_\mathcal{D} \equiv \int_\mathcal{D}d\vec{r}
  \sum_{i\in E_e}a_{g,i}b_i(\vec{r})\sum_{j\in E_e}b_{g,j}b_j(\vec{r}), \\
  \left(\bm{a},\bm{b}\right)_{\partial\mathcal{D}} \equiv
  \sum_{s\in\partial\mathcal{D}}\int_s d\vec{r}\sum_{i\in E_s}a_{g,i}b_i(\vec{r})\sum_{j\in E_s}
  b_{g,j}b_j(\vec{r}),
\end{gather}
%
involving integrals over the spatial domain to explicitly show the $S_N$ angular discretization.

Eqs.\ \ref{eq:time-derivative} to \ref{eq:reflecting-bc} present the terms from the \gls{SAAF}
$S_N$ equations in operator notation (Eq.\ \ref{eq:saaf-vt}) and their expanded forms
with explicit summations by neutron energy group and direction of neutron travel.

\noindent Time derivative term:
%
\begin{gather}
  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,
  \frac{\partial}{\partial t}\left(\frac{\bm{\Psi}}{\bm{v}}\right)\right) =
  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\Psi^*_{g,d},
  \frac{\Psi_{g,d}}{v_g}\right)_\mathcal{D} \label{eq:time-derivative}
\end{gather}
%
Streaming term:
%
\begin{gather}
  \left(\mathbb{L}_1\bm{\Psi}^*,
  \left(\bm{\tau}\mathbb{L}_1-\mathbb{I}+\bm{\tau}\mathbb{L}_2\right)\bm{\Psi}\right) =
  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},\tau_g\hat{\Omega}
  \cdot\nabla\Psi_{g,d}-(1-\tau_g\Sigma_{t,g})\Psi_{g,d}\right)_\mathcal{D}
\end{gather}
%
Collision term:
%
\begin{gather}
  \left(\mathbb{L}_2\bm{\Psi}^*,\bm{\Psi}\right) =
  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d},\Sigma_{t,g}\Psi_{g,d}\right)_\mathcal{D}
\end{gather}
%
Scattering term:
%
\begin{multline}
  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{S}\bm{\Psi}\right) = \\
  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
  \sum^G_{g'=1}\sum^L_{l=0}\Sigma^{g'\rightarrow g}_{s,l}\sum^l_{m=-l}
  \frac{2l+1}{w}Y_{l,m}(\hat{\Omega}_d)\phi_{g',l,m}\right)_\mathcal{D}
\end{multline}
%
Prompt fission source term:
%
\begin{gather}
  \left(\left(\mathbb{I}+\bm{\tau}\mathbb{L}_1\right)\bm{\Psi}^*,\mathbb{F}\bm{\Psi}\right) =
  \sum^G_{g=1}\sum^{N_d}_{d=1}w_d\left(\Psi^*_{g,d}+\tau_g\hat{\Omega}_d\cdot\nabla\Psi^*_{g,d},
  \frac{1}{w}\bar{\chi}_g\sum^G_{g'=1}\nu\Sigma_{f,g'}\phi_{g'}\right)_\mathcal{D}
\end{gather}
%
where $\bar{\chi}_g=\chi_{p,g}\left(1-\beta\right)$ for time-dependent problems with \glspl{DNP}
and $\bar{\chi}_g=\chi_{g}$ for steady-state problems without \glspl{DNP}.

@@

Regardless of the differences in implementations amongst the methods discussed here,
diffusion corrections applied through Eq.\ \ref{eq:saaf} were
found to be effective in enabling the neutron diffusion method to accurately reproduce reference
flux solutions \cite{gross_comprehensive_2023, pounders_diffusion_2009}.
However, a significant challenge for diffusion corrections
involves their resolution near neutron flux peaks. As observed in Eq.\ \ref{eq:svdc}, the neutron
current and flux gradient values do not necessarily reach zero at the same points in space,
resulting in diffusion coefficient values tending to positive or negative infinity when the
flux gradient is close to zero (see Figure \ref{fig:3b-svdc}. Larger mesh sizes may help avoid this
issue at the cost of worsening flux accuracy in regions with
steep, non-linear fluxes, such as near control rods. For the Ronen method, Gross et al.\
\cite{gross_comprehensive_2023} applied a numerical fix by avoiding correction
calculations wherever the flux gradient values fall below a certain threshold.

@@

equations. For this work, we adopt drift term derivations developed by Wang et al.\
\cite{wang_diffusion_2014, wang_rattlesnake_2018} specifically for the \gls{SAAF} $S_N$ equations
with void treatment. Their work uses drift terms for \gls{NDA}, while this work employs drift terms
for the hybrid method. We start by integrating Eq.\ \ref{eq:saaf-vt}

@@

In this discussion thus far, I have yet to establish the cutoff boundary for $V_1'$ within $V_1$
opposite to $\partial V_1$. A natural/intuitive criterion for the location of the cutoff boundary
would be where the components of the drift correction variable $\vec{D}_g$ become zero (i.e.,
where the components change signs). The reasons for this criterion are twofold. Firstly, at
points where the $\vec{D}_g$ components are zero, the flux is approximately isotropic along the
axes corresponding to the components because value of $\vec{D}_g$ represents how much correction a
standard neutron diffusion model requires. Secondly, this choice preserves the smoothness of the
neutron flux and flux gradient.
%
\begin{figure}[b]
	\centering
	\includegraphics[width=.75\columnwidth]{case-3b-geometry}
	\caption{1-D geometry for Case 3b.}
	\label{fig:3b-geometry}
\end{figure}
